### 2022-03-10 docs/papers/elements/abstract.tex
docs(paper): initialize paper in LaTeX, prepared to refactor it

%! Abstract
Increasing competition for water is challenging management institutions of large river basins around the world.
Institutions that successfully support sustainable water resource use are structurally well-aligned with water provisioning and social-ecological demands.
However, what constitutes a well-aligned institution in this context is poorly understood.
We analyzed institutional shifts in water resource allocation, exploiting two quasi-natural experiments of the Yellow River Basin, China.
Using a synthetic control identification strategy, we analyzed an institution that was intended to prevent water overuse but instead resulted in a social-ecological structural mismatch, leading to a 164\% increase beyond the expected total water use.
We applied a novel economic model that suggested that this “sprint effect” (i.e., rapid competing in resource use) and the resulting ecological crisis were driven by incentive distortions in water allocation bargaining, which in turn, led to each province trying to pre-emptively increase their water quotas for long-term gain.
Our analysis highlights the need to carefully evaluate institutional fit in the management of water in large river basins, particularly to avoid the possibility of incentive distortion.