### 2022-03-10 docs/papers/elements/info.tex

%! Author = songshgeo
%! Date = 2022/3/10

Funding was provided by the National Natural Science Foundation of China (CN) (Grant Nos. NSFC 42041007).

[Competing Interests] The authors declare that they have no competing interests.

[Correspondence] Correspondence and requests for materials should be addressed to Shuai Wang.~(email: shuaiwang@bnu.edu.cn).

[Author contributions]Shuai Wang and Bojie Fu designed this research, Shuang Song performed the research and analyzed data, and Huiyu Wen designed the economic model. Shuang Song and Huiyu Wen wrote the paper, and Graeme S. Cumming revised the manuscript and offered important advice.

[Code availability] All codes and datasets used in this research are accessible at https://github.com/SongshGeo/sochyd-transboundary-HESS. They will be made open source after this project is completed.