### 2022-03-20 docs/papers/tex/previous_appendix.tex

\documentclass{article}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[a4paper, total={7in, 10in}]{geometry}
\usepackage{amsmath}

\setstretch{1.523}

\newtheorem{ass}{Assumption}
\newtheorem{case}{Case}
%% make sure you have the nature.cls and naturemag.bst files where
%% LaTeX can find them

\bibliographystyle{naturemag}

\title{Supplementary Materials for

\input{../title.tex}}

%% Notice placement of commas and superscripts and use of &
%% in the author list

% \author{Shuang Song$^{1,2}$, Huiyu Wen$^3$, *Shuai Wang$^{1,2}$, Graeme S. Cumming$^{4}$ \& Bojie Fu$^{1, 2}$}

\begin{document}

\maketitle
\input{/Users/songshgeo/Documents/Pycharm/WAInstitution_YRB_2021/docs/papers/authors.md}

\begin{abstract}
    % 这个补充材料以下述逻辑组织而成：
    The supplementary material is organized logically as follows:
    % 展示流域水资源分配制度在全球的推行情况并详细介绍黄河流域的制度变化，旨在说明黄河流域为什么是一个独一无二的准自然实验。
    (1) Show the implementation of the water resource allocation system in the basin around the world and introduce the institutional changes in the Yellow River Basin in detail, aiming to explain why the Yellow River Basin is a unique quasi-natural experiment.
    % 展示定量分析方法的技术路线图，并详细介绍定量分析方法中的细节部分。
    (2) Show the technical roadmap of quantitative analysis method, and introduce the details of quantitative analysis method in detail.
    % 基于我们建立的一般经济模型，对一些可能的情景进行进一步分析。
    (3) Further extensions based on our general economic model.
\end{abstract}

\section*{S1. Detailed introduction of the institutions}
\input{../appendix/institution_context.tex}

\section*{S2. The technical roadmap and details of quantitative analyzing}
\input{../appendix/method_details.tex}

\section*{S3. Further analysis regarding the general economic model}
\input{../appendix/theoretical_model.tex}

\bibliography{WAInstitution_YRB_2021}
\end{document}