### 2022-05-20 docs/papers/02_appendix/02_method_details.tex

%! Author = songshgeo
%! Date = 2022/3/19

There are two approaches in validity testing of the DSC: (1) comparing the post- and pre- period reconstructions, (2) testing robustness through placebo analysis.

A good synthetic control will reconstruct (a) the outcome of the treated unit and (b) the covariates of the treated unit over the pre-treatment period. The plots and the Root Mean Square Percentage Error (RMSPE) help us evaluate (a), this table is meant to evaluate (b). If the synthetic control has good fit, the following things should be true:

Each row of the first two columns should be approximately equal. This means the synthetic control has reconstructed the treated unit values.

Placebo analysis is another standard approach to.