### 2022-07-03 docs/papers/01_manuscript/manuscript.tex

\documentclass[default, sn-standardnature]{sn-jnl}
\usepackage{amsmath,graphicx,array}
\usepackage{dcolumn,soul}%
\let\openbox\relax
\usepackage{amsthm}
\usepackage[figuresright]{rotating}%
\usepackage{algorithm, algorithmicx, algpseudocode}
\usepackage{listings}%
\usepackage{hyperref}
\newgeometry{left=2cm, right=2cm}
\linespread{1.5}
% \usepackage[nofiglist,notablist]{endfloat}
\makeatletter
\def\uns{\ifmmode\,\else$\,$\fi}%
\newtheorem{defn}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{ass}{Assumption}
\newtheorem{prop}{Proposition}
\newtheorem{fig}{Fig.}
\newtheorem{case}{Case}
\newtheorem{case_appendix}{Case}
\newtheorem{example}{Example}[section]
\renewcommand{\proofname}{\textbf{Proof}}
\newtheorem{property}{Property}
\newtheorem{remark}{Remark}
%\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{float}
\usepackage{multirow}
\usepackage{lineno}
\setenumerate{itemsep=0pt,partopsep=0pt,parsep=\parskip,topsep=0pt}
\setitemize{itemsep=0pt,partopsep=0pt,parsep=\parskip,topsep=5pt}
\setdescription{itemsep=0pt,partopsep=0pt,parsep=\parskip,topsep=5pt}

\usepackage{booktabs}
\usepackage{diagbox}
%%
\makeatother

% \jvol{XX}
% \jnum{X}
% \jyear{Year}
% \doi{10.1093/nsr/XXXX}
% \received{XX XX Year}
% \revised{XX XX Year}
% \accepted{XX XX Year}

\markboth{One, Two, and Three}{One, Two, and Three}
\graphicspath{{../../../figs/}}

\begin{document}
\normalsize
% \dhead{RESEARCH ARTICLE}
% \subhead{EARTH SCIENCES}

\bibliographystyle{../nsr}

\title{Approaching social-ecological matches of river basin systems for sustainability}  %! Notice Input

\author{Shuang Song$^{1}$}
\author{Huiyu Wen$^2$}
\author{*Shuai Wang$^{1}$}
\author{Xutong Wu$^{1}$}

\author{Graeme S. Cumming$^{3}$}
\author{Bojie Fu$^{1}$}

\affil{$^1$State Key Laboratory of Earth Surface Processes and Resource Ecology,
     Faculty of Geographical Science,
     Beijing Normal University,
     Beijing 100875,
     P.R. China}

% \affil{$^2$Institute of Land Surface System and Sustainability,
%      Faculty of Geographical Science,
%      Beijing Normal University,
%      Beijing 100875,
%      P.R. China}

\affil{$^2$School of Finance,
     Renmin University of China,
     Beijing 100875,
     P.R. China}

\affil{$^3$ARC Centre of Excellence for Coral Reef Studies,
     James Cook University,
     Townsville 4811,
     QLD, Australia}

% \affil{$^4$The research for this article was financed by the National Natural Science Foundation of China (CN) (Grant Nos. NSFC 42041007). A supplementary online appendix is available with this article at the \em{National Science Review} website.}

% \authornote{\textbf{Corresponding authors.} Email: shuaiwang@bnu.edu.cn}
%\authornote{Shuang Song and Huiyu Wen equally contributed to this work.}
\renewcommand\linenumberfont{\normalfont\bfseries\small}
\linenumbers
\abstract{
     Increasing competition for water is leading to depletion of freshwater globally and calls for efficient institutions in water governance.
     Alignments of social-ecological system (SES) structures are a crucial approach to institutional matches, but an understanding of its causal links and underlying processes are still weaknesses.
     To fill these knowledge gaps, we select the Yellow River Basin (YRB) in China as a typical case study to quantitatively measure the effects of SES structures changing because it is one of the most anthropogenically altered large river basins.
     Under different water allocation institutions, its streamflow was first overdrawn, then dried up, and finally has been successfully restored.
     We focused on two institutional shifts, the Water Allocation Scheme that began in 1987 (87-WAS) and the Unified Basinal Regulation that took over in 1998 (98-UBR), which re-framed different SES structures.
     We conduct counterfactual identification on the effect of these institutional shifts, our results suggested that during the decade following the introduction of the 87-WAS, observed water use of the YRB increased by $8.57\%$ more than expected, while 98-UBR ultimately decreased total water use.
     Furthermore, these heterogeneous effects of water use responses to SES structures aligned with our further theoretical marginal benefits analysis, supporting the hypothesis that SES structural changes played a vital role in sustainable water use.
     This quasi-natural experiment on the YRB offers profound insights into the links between SESs structures and outcomes, suggesting that fragmented ecological units linked to separated social actors should be avoided for sustainability.
     }  %! Notice input

\keywords{Yellow River, water use, water governance, social-ecological system, institutional fit}  %! Notice input
\maketitle

\section{INTRODUCTION}\label{sec:introduction}
\input{01_introduction.tex}

\section{RESULTS}\label{sec:results}
\input{02_structures.tex}
\input{03_results.tex}

\section{DISCUSSION}\label{sec:discussion}
\input{04_discussion.tex}

\section{CONCLUSION}\label{sec:conclusion}
\input{05_conclusion.tex}

\section{MATERIALS AND METHODS}\label{sec:methods}
\input{06_methods.tex}

\bibliography{../mybib}
\bibliographystyle{sn-standardnature}
\label{bib}

%%%%%%%% -----  02_appendix -------- %%%%%%%%%%
\newpage
\appendix
\label{appendix}
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\appendixname}{Appendix~\Alph{section}}
\setcounter{section}{0}

\section{Appendix~A: Contexts of institutional shifts}\label{secS1}
\renewcommand{\thefigure}{A\arabic{figure}}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}
\input{../02_appendix/01_institution_context.tex}

\section{Appendix~B: Robustness of DSC method}\label{secS2}
\renewcommand{\thefigure}{B\arabic{figure}}
\renewcommand{\thetable}{B\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}
\input{../02_appendix/02_method_details.tex}

% \section{S2: Methods in details}\label{secS3}
% \input{../02_appendix/03_results_appendix.tex}

\section{Appendix~C: Optimization model for water use}\label{secS4}
\renewcommand{\thefigure}{C\arabic{figure}}
\renewcommand{\thetable}{C\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}
\input{../02_appendix/04_theoretical_model.tex}

\section{Appendix~D: Model extensions}\label{secS5}
\renewcommand{\thefigure}{D\arabic{figure}}
\renewcommand{\thetable}{D\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}
\input{../02_appendix/05_model_extensions.tex}

%% Put the bibliography here, most people will use BiBTeX in
%% which case the environment below should be replaced with
%% the \bibliography{} command.
% \newpage
% \newpage
% \bibliography{../mybib}
% \bibliographystyle{sn-standardnature}
% \label{bib}
\end{document}