### 2023-04-22 docs/papers/01_manuscript/manuscript.tex

% chktex-file 46
\documentclass[preprint, 12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{amsmath, graphicx, array}
\usepackage{dcolumn, soul}%
\let\openbox\relax
\usepackage{amsthm}
\usepackage[figuresright]{rotating}%
\usepackage{algorithm, algorithmicx, algpseudocode}
\usepackage{listings}%
\usepackage{hyperref}
\usepackage{geometry}

\geometry{a4paper,left=2.0cm,right=2.0cm,top=2.5cm,bottom=2.5cm}
% 行距
\linespread{1.5}
% \usepackage[nofiglist,notablist]{endfloat}
% \def\uns{\ifmmode\,\else$\,$\fi}%
\newtheorem{defn}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{ass}{Assumption}
\newtheorem{prop}{Proposition}
\newtheorem{fig}{Fig.}
\newtheorem{case}{Case}
\newtheorem{case_appendix}{Case}
\newtheorem{example}{Example}[section]
\renewcommand{\proofname}{\textbf{Proof}}
\newtheorem{property}{Property}
\newtheorem{remark}{Remark}
%\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{float}
\usepackage{multirow}
% \usepackage{lineno}
\usepackage{booktabs}
\usepackage{diagbox}
%%

% \jvol{XX}
% \jnum{X}
% \jyear{Year}
% \doi{10.1093/nsr/XXXX}
% \received{XX XX Year}
% \revised{XX XX Year}
% \accepted{XX XX Year}

% \markboth{One, Two, and Three}{One, Two, and Three}
\graphicspath{{../../../figs/}}
\journal{Jounral of Hydrology}

\begin{document}
\begin{frontmatter}
% \normalsize
% \dhead{RESEARCH ARTICLE}
% \subhead{EARTH SCIENCES}

% \bibliographystyle{../nsr}
% 社会水文学专刊地址
% https://www.sciencedirect.com/journal/journal-of-hydrology/about/call-for-papers#grounded-sociohydrology
\title{Achieving Effective River Basin Governance through Institutional Shifts: Examining the Yellow River Basin's Social-Ecological Systems in China}

\author[inst1]{Shuang Song}
\author[inst2]{Huiyu Wen}
\author[inst1]{*Shuai Wang}
\author[inst1]{Xutong Wu}

\author[inst3]{Graeme S. Cumming}
\author[inst1]{Bojie Fu}

\affiliation[inst1]{
     State Key Laboratory of Earth Surface Processes and Resource Ecology,
     Faculty of Geographical Science,
     Beijing Normal University,
     Beijing 100875,
     P.R. China}

% \affil{stitute of Land Surface System and Sustainability,
%      Faculty of Geographical Science,
%      Beijing Normal University,
%      Beijing 100875,
%      P.R. China}

\affiliation[inst2]{School of Finance,
     Renmin University of China,
     Beijing 100875,
     P.R. China}

\affiliation[inst3]{
     ARC Centre of Excellence for Coral Reef Studies,
     James Cook University,
     Townsville 4811,
     QLD, Australia}

% \affil{e research for this article was financed by the National Natural Science Foundation of China (CN) (Grant Nos. NSFC 42041007). A supplementary online appendix is available with this article at the \em{National Science Review} website.}

% \authornote{\textbf{Corresponding authors.} Email: shuaiwang@bnu.edu.cn}
%\authornote{Shuang Song and Huiyu Wen equally contributed to this work.}
% \renewcommand\linenumberfont{\normalfont\bfseries\small}
% \linenumbers
\begin{abstract}
     Global freshwater depletion necessitates efficient water governance that accounts for social-ecological system (SES) structures. To improve our understanding of the relationships between SES structures and water governance, we investigated the Yellow River Basin (YRB) in China, which has experienced significant anthropogenic alterations. We analyzed two institutional shifts, the 1987 Water Allocation Scheme (87-WAS) and the 1998 Unified Basinal Regulation (98-UBR), that re-framed SES structures. Our counterfactual identification showed that water use in the YRB increased by 5.75\% more than expected following the 87-WAS, while the 98-UBR reduced total water use. These heterogeneous effects align with our marginal benefits analysis, confirming the importance of well-designed SES structures in water governance. Our quasi-natural experiment on the YRB highlights the need to avoid fragmented ecological units and separated social actors for sustainable water governance, while also emphasizing the significance of understanding the underlying processes and relationships in SES structures.
\end{abstract}

%%Graphical abstract
% \begin{graphicalabstract}
%      \includegraphics{outputs/main_results2.pdf}
% \end{graphicalabstract}

%%Research highlights
% \begin{highlights}
%      \item Research highlight 1
%      \item Research highlight 2
% \end{highlights}

\begin{keyword}
     water use~\sep~water governance~\sep~social-ecological system~\sep~institutions~\sep~Yellow River
\end{keyword}

\end{frontmatter}

\section{INTRODUCTION}\label{sec:introduction}
\input{01_introduction.tex}

\section{MATERIALS AND METHODS}\label{sec:methods}
\input{06_methods.tex}

\section{RESULTS}\label{sec:results}
\input{02_structures.tex}
\input{03_results.tex}

\section{DISCUSSION}\label{sec:discussion}
\input{04_discussion.tex}

\section{CONCLUSION}\label{sec:conclusion}
\input{05_conclusion.tex}

% Authors Contribution
% SW and BF designed this research. SS performed the study and analysed data. SS and HW wrote the paper. XW, GC, and HW revised and polished the manuscript and gave significant advice.

% Acknowledgments
% This research has been supported by the National Natural Science Foundation of China (grant no. 42041007) and the Fundamental Research Funds for the Central Universities

\bibliography{../mybib}
\bibliographystyle{elsarticle-num}\label{bib}

%%%%%%%% -----  02_appendix -------- %%%%%%%%%%
\newpage
\appendix\label{appendix}

\section{Contexts of institutional shifts}\label{secS1}
\renewcommand{\thefigure}{A\arabic{figure}}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}
\input{../02_appendix/01_institution_context.tex}

\section{Robustness of DSC method}\label{secS2}
\renewcommand{\thefigure}{B\arabic{figure}}
\renewcommand{\thetable}{B\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}
\input{../02_appendix/02_method_details.tex}

% \section{S2: Methods in details}\label{secS3}
% \input{../02_appendix/03_results_appendix.tex}

\section{Optimization model for water use}\label{secS4}
\renewcommand{\thefigure}{C\arabic{figure}}
\renewcommand{\thetable}{C\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}
\input{../02_appendix/04_theoretical_model.tex}

\section{Model extensions}\label{secS5}
\renewcommand{\thefigure}{D\arabic{figure}}
\renewcommand{\thetable}{D\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}
\input{../02_appendix/05_model_extensions.tex}

\end{document}