### 2023-05-05 docs/papers/01_manuscript/study_area.tex

The YRB, cradle of Chinese civilization, is located in north-central China and spans ten province-level regions whose socio-economic development heavily depends on water from the Yellow River.
As a semi-arid and arid region, the YRB's annual precipitation varies from about 100 to 1,000 mm and increases from the northwest to the southeast, while the annual pan evaporation varies from about 700 to 1,800 mm~\cite{wang2022e}.
Together, the YRB supports $35.63\%$ of China's irrigation and $30\%$ of its population while containing only $2.66\%$ of its water resources (data from \href{http://www.yrcc.gov.cn}{http://www.yrcc.gov.cn}, last access: \today).
Hence, over-withdrawing water from the Yellow River became an urgent concern when the river began to dry up in the early 1970s.
Among the policies proposed to address the problem, a series of water resource allocation institutions aimed to limit water use for each region with specific quotas, which were regarded as some of the most important solutions.
However, few attempts have been made to quantitatively assess how the YRB's water allocation scheme contributed to water governance, while other engineering solutions have been carefully evaluated~\cite{long2020}.
% 出现报错是因为最后的大写结尾LaTeX无法判断是缩写还是一个句子结束，这里是句子结束所以加 \\
% https://tex.stackexchange.com/questions/55105/when-should-i-use-intersentence-spacing

% Despite the essential reason for these institutions was the mismatch between the spatial and temporal distribution of water resources as well as social and economic water demands, the direct reason for their introduction was the depletion of the Yellow River.

The YRB was the first basin in China for which water resource allocation institutions were created, and institutional shifts can be traced through several regulating documents released by the Chinese government (at the national level):
(1) In $1980s$, proposed to develop a water resource allocation institution for the Yellow River~\cite{wang2019, wang2019a}.
(2) In $1987$, the Water Allocation Scheme was implemented (\href{http://www.gov.cn/zhengce/content/2011-03/30/content_3138.htm#}{http://www.mwr.gov.cn}, last access: \today).
(3) In $1998$, the Unified Basinal Regulation was implemented (\href{http://www.mwr.gov.cn/ztpd/2013ztbd/2013fxkh/fxkhswcbcs/cs/flfg/201304/t20130411_433489.html}{http://www.mwr.gov.cn}, last access: \today).
% 各省按要求编制新的黄河流域水资源规划，将水资源额度分配进一步细化。
(4) In $2008$, provinces were asked to draw up new water resources plans for the YRB to further refine water allocations~\cite{wang2019,wang2019a}.
(5) In $2021$, there was a call for redesigning the water allocation institution (\href{http://www.ccgp.gov.cn/cggg/zygg/gkzb/202107/t20210721_16591901.htm}{http://www.ccgp.gov.cn}, last access: \today).

Our study period therefore ranges from $1980$ (when water quotas were proposed) to $2008$, when a regulating system with quotas was fully established at basin, provincial, and district levels.
During this period, two significant institutional shifts can be analyzed using documents from $1987$ (87-WAS) and $1998$ (98-UBR), which split the study period into three sections: from $1980$ to $1987$ (before 87-WAS), from $1988$ to $1997$ (after 87-WAS and before 98-UBR), and from $1998$ to $2007$ (after 98-UBR).
% Those efforts led to ecological restoration of wetlands and the estuarine delta. Drying up has been avoided for over $20$ years, which is widely considered a substantial management achievement.