### 2023-05-06 docs/papers/01_manuscript/tab_rmse.tex

% Table generated by Excel2LaTeX from sheet 'Sheet1'
\begin{table}[!htbp]\footnotesize
	\centering
	\caption{Pre and post treatment root mean squared prediction error (RMSE) for YRB's provinces}\label{tab:DSC_summary}
	\begin{tabularx}{0.8\textwidth}{XXXXXXX}
	  \toprule
			& \multicolumn{3}{c}{87-WAS} & \multicolumn{3}{c}{98-UBR} \\
  \cmidrule{2-7}    province  & \multicolumn{1}{c}{post/pre} & To avg. & \multicolumn{1}{c}{sig.} & \multicolumn{1}{c}{post/pre} & To avg.   & \multicolumn{1}{c}{sig.} \\
	  \midrule
	  Qinghai & 5.26  & =     & FALSE & 5.89  & >     & TRUE \\
	  Gansu & 10.37  & >     & TRUE  & 9.55  & >     & TRUE \\
	  Ningxia & 5.81  & =     & FALSE & 6.83  & >     & TRUE \\
	  Inner Mongolia & 7.11  & >     & TRUE  & 1.60  & <     & TRUE \\
	  Shanxi & 1.72  & <     & TRUE  & 5.60  & >     & TRUE \\
	  Shaanxi & 3.05  & <     & TRUE  & 3.01  & >     & TRUE \\
	  Henan & 20.66  & >     & TRUE  & 1.18  & <     & TRUE \\
	  Shandong & 4.54  & =     & FALSE & 4.14  & >     & TRUE \\
	  \bottomrule
	  \end{tabularx}%
	\label{tab:addlabel}%
\end{table}%