### 2023-11-01 docs/papers/04_response_letter/source/editor.tex
docs(project): :memo: update R2 reviewer's comment

% chktex-file 46
\section*{Invited editorial team}\label{editor}

\subsection*{General comments}
\RC{} Dear Author,

% 现在有两位审稿人对修订后的版本进行了评估。他们都赞赏为考虑到对他们两人都审查过的初次提交的意见所做的工作。对论文的技术方面不太了解的人仍然对方法的解释很恼火。请仔细修改这方面，以确保期刊读者能理解你的论文
\RC*{} Two reviewers have now assessed the revised version. They both appreciate the work done to take in account the comments on the initial submission they had both reviewed. The one who is lcoser to the technical aspect of the paper is still quite annoyed though on the explanation of the method. Please revise carefully this aspect in order to ensure understandability of your paper by the journal readership

% TODO \AR{} I am grateful for the encouragement and the direction provided to enhance our manuscript. It has been a privilege to contribute to this special issue, and I believe that the revisions made align with the expectations and vision set forth by the editorial team.