### 2023-11-02 docs/papers/04_response_letter/source/editor.tex
fix(project): :memo: response letter

% chktex-file 46
\section*{Invited editorial team}\label{editor}

\subsection*{General comments}
\RC{} Dear Author,

% 现在有两位审稿人对修订后的版本进行了评估。他们都赞赏为考虑到对他们两人都审查过的初次提交的意见所做的工作。对论文的技术方面不太了解的人仍然对方法的解释很恼火。请仔细修改这方面，以确保期刊读者能理解你的论文
\RC*{} Two reviewers have now assessed the revised version. They both appreciate the work done to take in account the comments on the initial submission they had both reviewed. The one who is lcoser to the technical aspect of the paper is still quite annoyed though on the explanation of the method. Please revise carefully this aspect in order to ensure understandability of your paper by the journal readership

\AR{} We would like to express our sincere gratitude for the opportunity to revise our manuscript and for the constructive comments from the reviewers which have been instrumental in enhancing the quality of our paper.

\AR*{} Following the guidance provided, we have meticulously revised the manuscript, paying particular attention to the clarity of the methodological section as suggested. We have addressed all the concerns raised by Reviewer 1, including the accurate representation of mathematical equations and the logical sequence of our methodology.

\AR*{} We are confident that the changes made have significantly improved the manuscript, and we believe it is now ready for publication. We hope that the modifications meet the journal’s standards and look forward to the possibility of our study contributing to the journal's esteemed collection of research.

\AR*{} Thank you once again for your assistance and the reviewers for their essential insights.