### 2025-02-11 docs/papers/04_response_letter/source/editor.tex

% chktex-file 46
\section*{Invited editorial team}\label{editor}

\subsection*{General comments}
\RC{} Dear Author,

% 之前的审稿人现在已经审查了第二次修订版本。第一个审稿人仍然要求具体的澄清，下面有很好的解释。请务必考虑一下，我希望这将是一个最终的修订。
\RC*{} Both previous reviewers have now reviewed this second revised version. The first reviewer is still asking for specific clarifications well explained below.

\RC*{} Please do consider them in what I hope will be an ultimate revision.

% 很感谢审稿人提出的宝贵建议，它们的确能够让读者更加易于理解我们应用的计量方法，我们已经完全遵照其建议作出了修改。
\AR{} We are grateful to the reviewers for their valuable suggestions, which really make it easier for readers to understand the measurement methods we apply, and we have made changes in full compliance with their suggestions.