### 2025-02-11 docs/papers/04_response_letter/source/reviewer_2.tex

\section{Reviewer \#2}\label{reviewer_2}

\subsection*{Overall comments}

\RC{} The authors have addressed my concerns. I have no further comments or suggestions on this paper. Thank you.

\AR{} We are immensely grateful for your assessment and pleased to hear that you have no further comments. Your feedback has been invaluable to the improvement of our manuscript. Thank you for your time and contribution to our work.